On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-10-10 Manuscript/manuscript.tex
responded to Prof Yan's comments:

\begin{table}
  \centering
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
  from the same athletes who competed at the 2022 and 2023 World Track and Field
  Championships. Nat vs Inter compares data from the same athletes who competed
  at the 2022 national-level championships and the 2022 World Track and Field
  Championships.}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic \\ 
   \midrule
   2019 vs 2022 & $2.8\cdot10^{-2}$ & $2.9\cdot10^{-3}$ \\
   2022 vs 2023 & $1.1\cdot10^{-2}$ & $3.1\cdot10^{-3}$ \\
   2022 National vs International & $2\cdot10^{-6}$ & $1.1\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Clusrankresults}
\end{table}

@@

The estimated parameters of the GLMM model from our analysis are shown in 
Table~\ref{tab:Gamma_parameters}.  These statistics reflect the inclusion of 2022
in our analysis and shows the three models we considered. $\sigma_h^2$ is the
\jy{read what you wrote. These are variances not sd. Do not start a sentence
  with a symbol (see my writing guidelines).}
standard deviation of the heat effect, $\sigma_v^2$ is the standard deviation of
the venue effect, $\alpha$ is the intercept, and $\phi$ is the dispersion
parameter of the gamma model.
The GLMM with both venue and heat random
effects was the best model as evidenced by a decrease in the AIC compared to the
other models with only one random effect. 
Thus, results from the GLMM with both venue and heat random effects
will be presented throughout the rest
of the paper.

@@

\begin{table}
  \centering
  \caption{AIC and estimated parameters for the three GLMMs fitted to the ???
    data.}
  \label{tab:Gamma_parameters}
  \begin{tabular}{c c c c c c}
      \toprule
      Model & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\sqrt{\phi}$ \\
      \midrule
      Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
      Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
      Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
      \bottomrule
  \end{tabular}
\end{table}

Two aims of this section are to determine how extreme a reaction time below
0.1 seconds is, and to determine how the exclusion of 2022 could alter the results.
Using the gamma function with link set to log, the probability of the reaction 
time being less than 0.1 seconds can be calculated in R using the methods 
described in Section~\ref{sec:glmm}. A simulation size of $10,000,000$ was 
chosen to estimate the probability of observing a reaction time below 
two thresholds: 0.1 and 0.08. When including 2022, a probability
of $4.34\cdot10^{-3}$ corresponds to roughly 1 in every 230 starts
being below 0.1 seconds.  When we exclude 2022, the probability of observing a
reaction time below 0.1 seconds decreased to $3.54\cdot10^{-3}$, or 1 in every
283 starts.

@@

Table~\ref{tab:Sim_probability} shows that changing the reaction time 
barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
observing a reaction time that would break the barrier. If World Athletics were
to change the reaction time the way that \citet{komi2009iaaf} suggests, the 
probability would be so much lower that athletes would likely not
have a case to make when arguing against a disqualification.
\jy{discuss the differences between with/without 2022 data; comment on the
  pecularity of the 2022 data (as suggested by the rank based method); In a
  longer run, it may not be as strange as it looks at the first time.}

@@

\jy{be consistent in using scienfit format of numbers; 0.0001 vs $10^{-4}$}

We can use the same model but alter our interpretation slightly to select a 
reaction time barrier given the probability of observing a time less than the 
barrier.  When considering all of the data, if we want a 0.1\% chance of 
observing an illegally fast reaction time, the reaction
time barrier should be 0.092 seconds, as shown in Table~\ref{tab:Sim_time}.  We
can repeat this calculation for various probability levels,
for example a probability of 0.0001 (one disqualification for every 10,000 starts)
results in a reaction time barrier of 0.083 seconds.  As the exclusion of 2022
from the data resulted in a lower probability of observing a low reaction time,
the reaction time barrier will be less lenient when calculating the barrier
based on probability.

>>>>

\begin{table}
  \centering
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
  from the same athletes who competed at the 2022 and 2023 World Track and Field
  Championships. National vs International compares data from the same athletes
  who competed at 2022 national-level championships and the 2022 World Track and
  Field Championships.}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic \\ 
   \midrule
   2019 vs 2022 & $2.8\cdot10^{-2}$ & $2.9\cdot10^{-3}$ \\
   2022 vs 2023 & $1.1\cdot10^{-2}$ & $3.1\cdot10^{-3}$ \\
   2022 National vs International & $2\cdot10^{-6}$ & $1.1\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Clusrankresults}
\end{table}

@@

The estimated parameters of the GLMM model from our analysis are shown in 
Table~\ref{tab:Gamma_parameters}. These statistics reflect the inclusion of 2022
in our analysis and shows the three models we considered. The standard deviation
of the heat effect is $\sigma_h^2$, the standard deviation of the venue effect
is $\sigma_v^2$, the intercept is $\alpha$ and we report $\sqrt{\phi}$ as the
standard deviation of the dispersion parameter $\phi$
\jy{read what you wrote. These are variances not sd. Do not start a sentence
  with a symbol (see my writing guidelines).}
\of{Resolved}
The GLMM with both venue and heat random
effects was the best model as evidenced by a decrease in the AIC compared to the
other models with only one random effect. 
Thus, results from the GLMM with both venue and heat random effects
will be presented throughout the rest
of the paper.

@@

\begin{table}
  \centering
  \caption{AIC and estimated parameters for the three models fitted to all the
    data.}
  \label{tab:Gamma_parameters}
  \begin{tabular}{c c c c c c}
      \toprule
      Model & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\sqrt{\phi}$ \\
      \midrule
      Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
      Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
      Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
      \bottomrule
  \end{tabular}
\end{table}

Two aims of this section are to determine how extreme a reaction time below
0.1 seconds is, and to determine how the exclusion of 2022 could alter the results.
Using the gamma function with link set to log, the probability of the reaction 
time being less than 0.1 seconds can be calculated in R using the methods 
described in Section~\ref{sec:glmm}. A simulation size of $10,000,000$ was 
chosen to estimate the probability of observing a reaction time below 
two thresholds: 0.1 and 0.08.

@@

Table~\ref{tab:Sim_probability} shows that changing the reaction time 
barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
observing a reaction time that would break the barrier. If World Athletics were
to change the reaction time the way that \citet{komi2009iaaf} suggests, the 
probability would be so much lower that athletes would likely not
have a case to make when arguing against a disqualification. When including 2022,
a probability of $4.34\cdot10^{-3}$ corresponds to roughly 1 in every 230 starts
being below 0.1 seconds. When we exclude 2022, the probability of observing a
reaction time below 0.1 seconds decreased to $3.54\cdot10^{-3}$, or 1 in every
283 starts.  While this seems low, it is not unreasonable to think that certain
athletes with strong reaction times could begin approach the 0.1 second barrier,
especially when considering Allen's other reaction times: 0.124 and 0.101 at the
2022 World Track and Field Championships.  Another important conclusion from our
results is how excluding 2022 lowered the probability of observing a fast 
reaction time, suggesting that 2022 was peculiarly fast, which aligns with the
results presented in section~\ref{subsec:Results_Rank}.

\jy{discuss the differences between with/without 2022 data; comment on the
  pecularity of the 2022 data (as suggested by the rank based method); In a
  longer run, it may not be as strange as it looks at the first time.}
\of{Resolved}

@@

\jy{be consistent in using scienfit format of numbers; 0.0001 vs $10^{-4}$}
\of{In our meeting you said to convert scientific notation like $10^{-2}$ to 
0.01, and I agree I think 0.01 makes much more sense. So should I convert
everything to decimals or scientific notation?}

We can use the same model but alter our interpretation slightly to select a 
reaction time barrier given the probability of observing a time less than the 
barrier. When considering all the data, if we want a 0.1\% chance of 
observing an illegally fast reaction time, the reaction
time barrier should be 0.092 seconds, as shown in Table~\ref{tab:Sim_time}. We
can repeat this calculation for various probability levels,
for example a probability of 0.0001 (one disqualification for every 10,000 starts)
results in a reaction time barrier of 0.083 seconds. As the exclusion of 2022
from the data resulted in a lower probability of observing a low reaction time,
the reaction time barrier will be less lenient when calculating the barrier
based on probability.
